Astro 507; Problem Set 3

### 2022-02-23 pset3/solutions.tex
2a complete

\question{\textbf{1. COBE}}

\question{\textbf{2. Entropy}}

\question{\textbf{2. Entropy}}

>>>>

\question{1. \textbf{FIRAS}}

\answer{}

\question{2a. \textbf{Entropy}}
\answer{
    Let's derive the entropy for an ideal, non-relativistic Fermi gas (in terms of $V$, $z$ and $T$). For this I will follow Lecture 10. The entropy is defined as
    \begin{equation}
        S = \frac{U + PV - N \mu}{T},
    \end{equation}
    so we just need to use definitions of $U$, $P$, $N$ and $\mu$. The definitions of $\mu$ is trivial when using the fugacity
    \begin{equation}
        \mu = k_B T \ln z
    \end{equation}
    For the others, I refer to lecture 10 where we can find the definition of $n$ on slide 5
    \begin{equation}
        N = n V = \frac{2(2s + 1)}{\pi^{1/2} \lambda^3} V F_{1/2}(z)
    \end{equation}
    and the definition of $P$ on slide 7
    \begin{equation}
        P = \frac{4(2s + 1)}{3 \pi^{1/2}} k_B T \lambda^{-3} F_{3/2}(z)
    \end{equation}
    For the total energy, we don't have it from the slides but we can derive it in the same way using Fermi-Dirac integrals. It starts in the same way as $n$ but with an extra factor of $\epsilon$ in the numerator, then we use the substitution $w = \epsilon / k_B T$
    \begin{align}
        U &= (2s + 1) \frac{V}{h^3} \int \frac{\epsilon}{e^{\frac{\epsilon - \mu}{k_B T}} + 1} \dd{\va{p}} \\
        U &= \frac{4\pi (2s + 1)}{2 h^3} (2 m k_B T)^{3/2} k_B T \int_0^\infty \dd{w} \frac{w^{3/2}}{e^w z^{-1} + 1} \\
        U &= \frac{2(2s + 1)}{\pi^{1/2} \lambda^3} V k_B T F_{3/2}(z)
    \end{align}
    Now it's simply a matter of plugging all of these into the entropy expression.
    \begin{align}
        S &= \frac{U + PV - N \mu}{T} \\
          &= \frac{1}{T} \qty[ \frac{2(2s + 1)}{\pi^{1/2} \lambda^3} V k_B T F_{3/2}(z) + \frac{4(2s + 1)}{3 \pi^{1/2}} V k_B T \lambda^{-3} F_{3/2}(z) + \frac{2(2s + 1)}{\pi^{1/2} \lambda^3} V k_B T \ln z F_{1/2}(z)] \\
          &= \frac{2(2s + 1) V k_B T}{\pi^{1/2} \lambda^3} \qty[ F_{3/2}(z) + \frac{2}{3} F_{3/2}(z) + \ln z F_{1/2}(z)] \\
        \Aboxed{ S &= \frac{2(2s + 1) V k_B T}{\pi^{1/2} \lambda^3} \qty[ \frac{5}{3} F_{3/2}(z) + F_{1/2}(z) \ln z] }
    \end{align}
}

\question{2b. \textbf{Expanding pressure}}
\answer{
    \begin{align}
        \frac{P}{n k_B T} &= \frac{4(2s + 1)}{3 \pi^{1/2}} k_B T \lambda^{-3} F_{3/2}(z) \qty[\frac{2(2s + 1)}{\pi^{1/2} \lambda^3} k_B T F_{1/2}(z)]^{-1} \\
                          &= \frac{2 F_{3/2}(z)}{3 F_{1/2}(z)}
    \end{align}
    \todo{come back to this on Thursday}
}

\question{\textbf{3. Brown Dwarf}}

\answer{}
